Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-08-17 writeup/conStruct.tex
added fastStructure results

There have been many methods proposed to characterize populations,
including generating population phylogenies \citep{CavalliSforza1975, treemix},
$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
Like the method we present here, these are based on a statistical model,
but do not perform inference directly with a demographic model.
Each of these methods perform best in particular situations,
but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.

@@

We ran \texttt{conStruct} analyses on each of these simulated datasets, 
specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
then used cross-validation to compare the predictive performance of the models.
In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.

@@

\begin{figure}
	\centering
		\subcaptionbox{$K=2$\label{simK3_sp_pies:K2}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K2.pdf}}
		\subcaptionbox{$K=3$\label{simK3_sp_pies:K3}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K3.pdf}}
		\subcaptionbox{$K=4$\label{simK3_sp_pies:K4}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K4.pdf}}
		\subcaptionbox{$K=5$\label{simK3_sp_pies:K5}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K5.pdf}}
		\subcaptionbox{$K=6$\label{simK3_sp_pies:K6}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K6.pdf}}
		\subcaptionbox{$K=7$\label{simK3_sp_pies:K7}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K7.pdf}}
	\caption{
	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
    }\label{simK3_sp_pies}
\end{figure}

>>>>

There have been many methods proposed to characterize populations,
including generating population phylogenies \citep{CavalliSforza1975, treemix},
$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, fastStructure, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
Like the method we present here, these are based on a statistical model,
but do not perform inference directly with a demographic model.
Each of these methods perform best in particular situations,
but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.

@@

We ran \texttt{conStruct} analyses on each of these simulated datasets, 
specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
then used cross-validation to compare the predictive performance of the models.
In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
To confirm that results from our nonspatial model were qualitatively similar to those of 
other nonspatial model-based clustering algorithms, 
we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
inferring admixture under models with $K=2$ through 4 (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).

@@

\begin{figure}
	\centering
		\subcaptionbox{$K=2$\label{simK3_sp_pies:K2}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K2.pdf}}
		\subcaptionbox{$K=3$\label{simK3_sp_pies:K3}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K3.pdf}}
		\subcaptionbox{$K=4$\label{simK3_sp_pies:K4}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K4.pdf}}
		\subcaptionbox{$K=5$\label{simK3_sp_pies:K5}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K5.pdf}}
		\subcaptionbox{$K=6$\label{simK3_sp_pies:K6}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K6.pdf}}
		\subcaptionbox{$K=7$\label{simK3_sp_pies:K7}}
			{\includegraphics[width=1.8in,height=1.8in]{figs/sims/simK3_sp_pies_K7.pdf}}
	\caption{
	Map of admixture proportions estimated using a spatial model for $K=2$ through 7.
	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
    }\label{simK3_sp_pies}
\end{figure}
\clearpage

\begin{figure}
	\centering
		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK1_pies.pdf}}
	\caption{
	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
	The data were simulated using one cluster with nearest-neighbor symmetric migration between demes.
    }\label{fastStr_simK1}
\end{figure}

\begin{figure}
	\centering
		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK2_pies.pdf}}
	\caption{
	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
	The data were simulated using two clusters with nearest-neighbor symmetric migration between demes.
    }\label{fastStr_simK2}
\end{figure}

\begin{figure}
	\centering
		{\includegraphics[width=6in,height=2.25in]{figs/fastStr/fastStr_simK3_pies.pdf}}
	\caption{
	Map of admixture proportions estimated using fastSTRUCTURE \citep{fastStructure} for $K=2$ through 4.
	The data were simulated using three clusters with nearest-neighbor symmetric migration between demes.
    }\label{fastStr_simK3}
\end{figure}
